---
title: "EDLD 610 Final Project"
subtitle: "Winter 2020"
author: "Jim Wright"
output:
  xaringan::moon_reader:
    css: ["default", "uo", "uo-fonts"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(flexdashboard)
library(tidyverse)
library(rio)
library(here)
library(colorblindr)
library(gghighlight)
library(forcats)
library(ggrepel)
library(gt)
library(knitr)
library(kableExtra)
library(reactable)
library(plotly)

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
               fig.width = 13)

theme_set(theme_minimal(base_size = 8))


all_ages <- import(here("data", "all-ages.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 


grad_students <- import(here("data", "grad-students.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

majors_lists <- import(here("data", "majors-list.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

recent_grads <- import(here("data", "recent-grads.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

women_stem <- import(here("data", "women-stem.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 
```


# Data Description 

* Data from FiveThirtyEight GitHub page  (https://github.com/fivethirtyeight/data/tree/master/college-majors)
* Consists of five data sets from the American Community Survey 2010-2012 Public Use Microdata Series
* 5 total data sets with information on basic earnings and labor information 
* 2 primary data sets used for project: recent_grads.csv and women-stem.csv
* Both arranged in order of median salary for reported college major on the survey 


---
# Plot 1

* Goal is to communicate median salary per either college major or major category variables
* 5 total versions

---

# Version 1

* College major variable contained too many values for a bar graph 
* College major_category was easier to plot (16 values)
* Key point: Engineering major category has highest median salary by a A LOT

```{r plot 1 data cleaning, include=FALSE}

recent_grads %>% 
  select(major, median, p25th, p75th)

```

```{r plot 1 version 1, fig.height=6}
ggplot(recent_grads, aes(major_category, median)) +
  geom_col() +
  coord_flip()
```

---

# Version 2

Added some color to the major category of Education 

```{r plot 1, version 2}

ggplot(recent_grads, aes(fct_reorder(major_category, median), median)) +
  geom_col() +
  geom_col(fill = "cornflowerblue",
  data = filter(recent_grads, major_category == "Education")) +  
  coord_flip()

```

---

# Plot 1 Versions 3 and 4

* Wanted to explore alternative plot options
* Plot 3 - histogram
* Plot 4 - density plot

## Version 3

```{r plot 1, version 3, fig.height=5.5}
ggplot(recent_grads, aes(median)) +
  geom_histogram(fill = "#56B4E9",
                 color = "black", 
                 alpha = 0.9,
                 bins = 25) +
  theme_minimal(base_size = 15) +
  scale_x_continuous("Median Salary of College Majors", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Median Salary of College Majors",
       y = "Total",
       title = "Distribution of Median Salaries \nper College Major") +
  theme(plot.title = element_text(hjust = 0.5))

```


---
# Version 4

```{r plot 1, version 4}
ggplot(recent_grads, aes(median, major_category)) +
  ggridges::geom_density_ridges(color = "white",
                                fill = "#A9E5C5") +
  theme_minimal(base_size = 10) +
  scale_x_continuous("Median Salary of College Majors", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Median Salary Distribution", 
       y = "Major Category",
       title = "Distribution of Median Salaries \nby Major Category") +
  theme(plot.title = element_text(hjust = 0.5))
```

---

# Final Version

* Returned to the bar graph 
* Sliced the top 20 median salaries by college major to use that variable 

```{r plot 1 final version data cleaning, include=FALSE}
median_top_20 <- recent_grads %>% 
  arrange(desc(median)) %>% 
  slice(1:20)
```

```{r plot 1, final version}
ggplot(median_top_20, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  geom_text(aes(major, median, label = scales::dollar(median/1000)),
            hjust = 1.1,
            nudge_y = 0.02,
            size = 3,
            color = "white") +
  scale_fill_brewer(palette = "Dark2") +
  coord_flip() +
  scale_y_continuous("Median Salary", 
                     labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "College Major",
       y = "Median Salary in Thousands",
       fill = "Major Category",
       title = "Top 20 Median Salaries \nby College Major") 
```

---
# Plot 2

.pull-left[

## Details
* Goal is to communicate unemployment rates across college major and major category variables
* 3 total versions
]

.pull-right[

## Variables and Functions Used
* Continous variables representing male and female employment totals per major
* mutate() function was used to make these values percentages per major
* pivot_longer() used to create one variable with individual values for percentage of males and percentage of females
]

---
# Version 1
```{r plot 2 data cleaning, include=FALSE}

percentages <- recent_grads %>% 
  mutate(percent_men = men/total, 
         percent_women = women/total, 
         total_jobs = college_jobs + non_college_jobs + low_wage_jobs, 
         percent_college_jobs = college_jobs/total_jobs,
         percent_non_college_jobs = non_college_jobs/total_jobs, 
         percent_low_wage_jobs = low_wage_jobs/total_jobs,
         unemployment_rate_percentage = unemployment_rate)

percents_tidy <- percentages %>% 
  pivot_longer(
    cols = c(22:23),
    names_to = "sex",
    values_to = "sex_percentages"
  ) 

percentages %>% 
  summarize(mean(median))

percentages %>% 
  summarize(mean(unemployment_rate_percentage))

plot_2_a <- percents_tidy %>% 
  filter(median >= 40000,
         unemployment_rate_percentage <= 6.50)

plot_2_b <- percents_tidy %>% 
  filter(median <= 40000,
         unemployment_rate_percentage >= 6.50)
```

```{r plot 2 version 1}
ggplot(plot_2_a, aes(unemployment_rate_percentage, median, color = major_category)) +
  geom_point(size = 2,
             alpha = 0.5) +
  scale_color_viridis_d(name = "Major Category") +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  scale_x_log10("Unemployment Rate Percentage", labels = scales::percent) +
  scale_y_log10("Median Salary", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Unemployment Rate Percentage",
       y = "Median Salary",
       title = "Comparison of Employment Rate to Median Salary",
       subtitle = "College Majors with Above Average Median Salary and Below Average Unemployment Rate") 

```

#### A little hard to interpret 


--- 

# Version 2

```{r plot 2 version 2}
ggplot(percents_tidy, aes(fct_reorder(major_category, unemployment_rate), unemployment_rate)) +
  geom_col(fill = "cornflowerblue",
           alpha = 0.7) +
  geom_text(aes(major_category, unemployment_rate, label = paste0(round(unemployment_rate), "%",
            size = 2))) +
  coord_flip() +
  scale_y_continuous("Unemployment Rate", labels = scales::percent) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "Major Category",
       title = "Unemployment Rates per Major Category") 

```

#### Could not get my text labels to work!

---

# Final Version 

```{r plot 2 final version data cleaning, include=FALSE}
plot_2_c <- percents_tidy %>% 
  select(major, major_category, unemployment_rate_percentage, percent_college_jobs) %>% 
  arrange(desc(unemployment_rate_percentage)) %>% 
  slice(1:20)
```

```{r plot 2 final version, fig.height=8}
plot_2_c %>% 
  distinct() %>% 
  ggplot(aes(unemployment_rate_percentage, percent_college_jobs, color = major_category))+
  geom_point(size = 2) +
  scale_color_viridis_d(name = "Major Category") +
  geom_text_repel(aes(label = major),
                  size = 3) +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  scale_x_log10("Unemployment Rate", labels = scales::percent) +
  scale_y_log10("Employment in College Major", labels = scales::percent) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Unemployment Rate",
       y = "Employment in College Major",
       title = "Comparison of Unemployment Rate \nto College Major Employment Rate",
       subtitle = "College Majors with the Highest Unemployment Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```


---
# Plot 3

.pull-left[

## Details
* Goal is to communicate employment rate of women in STEM related fields 
* 3 total versions and two additional tables 
* Learned that gt tables don't fit well on flexdashboards or slides
]

.pull-right[

## Variables and Functions Used
* major and major category
* used filter() to focus on engineering major category for versions 1 and 2
* final version included slice() function to identify top 10 and bottom 10 for women employment   rate
* then used bin_rows() to combine these objects for final plot
]

---
# Version 1
```{r plot 3 data cleaning, include=FALSE}

stem_percentages <- women_stem %>% 
  mutate(percent_men = men/total,
         percent_women = women/total)

stem_tidy <- stem_percentages %>% 
  pivot_longer(
    cols = c(10:11),
    names_to = "sex",
    values_to = "sex_percentages"
  )

engineering <- stem_percentages %>% 
  filter(major_category == "Engineering")

engineering_2 <- stem_tidy %>% 
  filter(major_category == "Engineering")

```

```{r plot 3 version 1}
ggplot(engineering, aes(fct_reorder(major, percent_women), percent_women)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous("Percentage of Women Employed", labels = scales::percent) 

```

--- 

# Version 2

```{r plot 3 version 2}
ggplot(engineering, aes(fct_reorder(major, percent_women), percent_women)) +
  geom_col(fill = "#0000FF",
          alpha = 0.7) +
  geom_text(aes(major, percent_women, label = scales::percent(percent_women, digits = 2)),
            nudge_y = -0.05,
            size = 3,
            color = "white") +
  coord_flip() +
  scale_y_continuous("Percentage of Women Employed", labels = scales::percent) 

```

---

# Final Version

```{r plot 3 final version data cleaning, include=FALSE}
top_10 <- stem_percentages %>% 
  group_by(major_category) %>% 
  top_n(2, percent_women)

bottom_10 <- stem_percentages %>% 
  group_by(major_category) %>% 
  top_n(-2, percent_women)
```

```{r plot 3 final version, fig.width=8, fig.height=6}
ggplot(stem_percentages, aes(percent_women, percent_men)) +
  geom_point(color = "gray80") +
  geom_point(color = "red", data = bind_rows(top_10, bottom_10)) +
  geom_text_repel(aes(label = major),
                  data = bind_rows(top_10, bottom_10),
                  size = 2) +
  facet_wrap(~major_category) +
  theme_minimal() +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5)) +
  scale_x_log10("Percent Women Employed", labels = scales::percent) +
  scale_y_log10("Percent Men Employed", labels = scales::percent) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5)) +
  labs(title = "Gender Employment Percentages in STEM Related Fields")

```

---

# Table Practice! 

## GT Table

```{r plot 3 gt table data cleaning, include=FALSE}
table <- bind_rows(top_10, bottom_10) %>% 
  select(major, median, percent_women, percent_men) 
```

```{r plot 3 gt table}
table %>% 
  gt() %>% 
  cols_label(major = "College Major",
             median = "Median Salary",
             percent_women = "Percent Women Employed",
             percent_men = "Percent Men Employed") %>% 
  cols_align(align = "left", columns = vars(major)) %>% 
  cols_align(align = "center", columns = vars(percent_men, percent_women)) %>% 
  tab_header(title = "STEM Major Statistics") %>% 
  fmt_percent(vars(percent_women, percent_men), decimals = 0) %>% 
  fmt_currency(vars(median), currency = "USD") %>% 
  cols_align(align = "left", columns = vars(major)) %>% 
  tab_source_note(source_note = md("Data from [fivethirtyeight](https://https://github.com/fivethirtyeight/data/tree/master/college-majors)"))
```

---

# Reactable Table 

```{r plot 3, reactable table data cleaning, include=FALSE}

table_react <- bind_rows(top_10, bottom_10) %>% 
  select(major_category, major, median, percent_women, percent_men) %>% 
  rename("Major Category" = major_category,
         Major = major,
         "Median Salary" = median,
         "Percent Women Employed" = percent_women,
         "Percent Men Employed" = percent_men)

```

```{r plot 3, reactable table}

reactable(table_react, columns = list(
  "Median Salary" = colDef(format = colFormat(prefix = "$", separators = TRUE, digits = 2)),
  "Percent Women Employed" = colDef(format = colFormat(percent = TRUE, digits = 2)),
  "Percent Men Employed" = colDef(format = colFormat(percent = TRUE, digits = 2))
))

```

